### 2025-07-20 decision-making.qmd
20250720 - summary of prediction accuracy

Even though the crowd tends to be more accurate than individual forecasters, crowd-averaged projections (at least among experts) are not necessarily highly accurate.
In fantasy football, [DESCRIBE THEIR ACCURACY--BOTH FOR ALL PLAYERS AND WHEN SUBSETTING TO THE TOP X PLAYERS].
Nevertheless, individual sources tend to be even less accurate.
[DESCRIBE THEIR ACCURACY].

>>>>

Even though the crowd tends to be more accurate than individual forecasters [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>], crowd-averaged projections (at least among experts) are not necessarily highly accurate, as described in @sec-predictionAccuracyByPosition.
In fantasy football, crowd-averaged seasonal projections explain ~60–75% of the variance in fantasy points among offensive players; however, the percent of variance explained drops to ~30% when considering only those players with high projected or actual points—who are the players that are the most important to distinguish, because those are the players that you are most likely trying to decide between when drafting.
Nevertheless, individual sources tend to be even less accurate.
Individual projections sources tend to explain ~50–65% of the variance in fantasy points among offensive players; however, the percent of variance explained also drops considerably when considering only those players with high projected or actual points.

### 2025-07-20 evaluating-prediction-accuracy.qmd
20250720 - summary of prediction accuracy

Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.
In general, projections showed moderate accuracy in predicting players' fantasy performance.
Projections showed moderate accuracy in terms of [discrimination](#sec-discrimination), based on [$R^2$](#sec-coefficientOfDetermination).
However, projections were not highly accurate.
Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity form of [miscalibration](#sec-calibration).
As a reminder, the overextremity form of [miscalibration](#sec-calibration) is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.
Relatedly, crowd-averaged projections tend to be more accurate than individual sources of projections [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].

@@

@@

@@

@@

@@

On the whole, experts' predictions are inaccurate.
Experts' predictions from many different domains tend to be inaccurate, including political scientists [@Tetlock2017], physicians [@Koehler2002], clinical psychologists [@Oskamp1965], stock market traders and corporate financial officers [@Skala2008], seismologists' predictions of earthquakes [@Hough2016], economists' predictions about the economy [@Makridakis2009], lawyers [@Koehler2002], and business managers [@Russo1992].
Thus, I would not put much confidence in the predictions by fantasy football fundits.
The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in Section @sec-calibration.
Overextremity of experts' predictions reflects the [overprecision](#sec-cognitiveBiasesOverconfidence) type of [overconfidence bias](#sec-cognitiveBiasesOverconfidence).
The degree of confidence of a person's predictions is often not a good indicator of the accuracy of their predictions [and confidence and prediction accuracy are sometimes inversely associated; @Silver2012].
[Heuristics](#sec-heuristics) such as the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment), [cognitive biases](#sec-cognitiveBiases) such as [confirmation bias](#sec-cognitiveBiasesConfirmation) [@Hoch1985; @Koriat1980], [fallacies](#sec-fallacies) such as the [base rate fallacy](#sec-fallaciesBaseRate) [@Eddy1982; @Koehler2002] could contribute to overconfidence of predictions.
[Poorly calibrated](#sec-calibration) predictions are especially likely when the [base rate](#sec-baseRate) is very low (e.g., suicide) or when the [base rate](#sec-baseRate) is very high [@Koehler2002].

>>>>

Below are estimates of projection accuracy and [calibration plots](#sec-calibrationPlot) for [seasonal](#sec-accuracyProjectionsSeasonal) and [weekly](#sec-accuracyProjectionsWeekly) projections.

@@

In the data object, each projection source has a row for each player and season.
Thus, the $R^2$ may be higher for this aggregated object than it is for many of the individual sources.
To calculate the accuracy by source, we could use the following code:

```{r}
accuracy_by_source <- projectionsWithActuals_seasonal %>%
  filter(!is.na(data_src)) %>% 
  filter(data_src != "FantasyPros") %>% # remove crowd-averaged source; we compare to crowd average later
  group_by(data_src) %>%
  summarise(
    accuracy = list(
      accuracyOverall(
        predicted = raw_points,
        actual = fantasyPoints,
        dropUndefined = TRUE
      )
    ),
    .groups = "drop"
  )

accuracy_by_source
accuracy_by_source$accuracy
```

@@

We summarize the accuracy of crowd-averaged seasonal projections by position in Tables [-@tbl-seasonalCrowdAccuracyByPosition] and [-@tbl-seasonalCrowdAccuracyByPositionSubset].

@@

We summarize the accuracy of crowd-averaged seasonal projections by position in Tables [-@tbl-weeklyCrowdAccuracyByPosition] and [-@tbl-weeklyCrowdAccuracyByPositionSubset].

@@

In general, projections showed moderate accuracy in predicting players' fantasy performance.
Projections showed moderate accuracy in terms of [discrimination](#sec-discrimination), based on [$R^2$](#sec-coefficientOfDetermination).
However, projections were not highly accurate.
Projections showed some [miscalibration](#sec-calibrationIndices) with respect to actual fantasy points.
In general, projected showed a modest tendency to underpredict performance at lower levels of projected points and a moderate-to-strong tendency overpredict performance at higher levels of projected points, an example of the overextremity form of [miscalibration](#sec-calibration).
As a reminder, the overextremity form of [miscalibration](#sec-calibration) is an example of the overprecision form of [overconfidence](#sec-cognitiveBiasesOverconfidence).
That is, projection sources appear to be [overconfident](#sec-cognitiveBiasesOverconfidence) in their projections of player performance.

For seasonal projections, in terms of comparison by position, projections showed greatest discrimination ($R^2$) for Quarterbacks, followed by Running Backs/Tight Ends, and Wide Receivers.
Predictions for defensive players and kickers showed a negative $R^2$, indicating that projections would have been more accurate merely by predicting the mean of the actual values for every observation (i.e., [predicting from the base rate](#sec-predictingFromBaseRate)).
In terms of comparison by projection source, crowd-averaged projections showed greater accuracy than individual projection sources (i.e., "experts") in terms of discrimination ($R^2$) and calibration (MAE).
This was especially the case when examining individual sources separately (rather than in one combined object).

For weekly projections, in terms of comparison by position, projections showed greatest discrimination ($R^2$) for Running Backs, followed by Wide Receivers, Quarterbacks, and Tight Ends.
Predictions for defensive players and kickers showed a negative $R^2$, indicating that projections would have been more accurate merely by predicting the mean of the actual values for every observation (i.e., [predicting from the base rate](#sec-predictingFromBaseRate)).
In terms of comparison by projection source, crowd-averaged projections showed greater accuracy than individual projection sources in terms of discrimination ($R^2$) and calibration (MAE).

In summary, we found that, for both seasonal and weekly projections, crowd-averaged projections showed greater accuracy than individual projection sources (i.e., "experts"), consistent with prior findings [@Kartes2024; archived at <https://perma.cc/69F7-LLTN>; @Petersen2017; archived at <https://perma.cc/BG2W-ANUF>].
Nevertheless projections—including individual sources and crowd-averaged—were not highly accurate.
Projections tended not to explain more than about two-thirds of the variance in players' performance, and frequently explained well less than half of the variance, particularly when distinguishing between players with high (i.e, more than 100) projected or actual points—who are the players that are the most important to distinguish, because those are the players that you are most likely trying to decide between when drafting.
Thus, although projections for offensive positions (besides Kickers) do better than chance, there remains considerable room for improvement in predictive accuracy.
The modest accuracy of projections likely reflects that human behavior is multiply determined and hard to predict.

@@

On the whole, experts' predictions are inaccurate.
Experts' predictions from many different domains tend to be inaccurate, including political scientists [@Tetlock2017], physicians [@Koehler2002], clinical psychologists [@Oskamp1965], stock market traders and corporate financial officers [@Skala2008], seismologists' predictions of earthquakes [@Hough2016], economists' predictions about the economy [@Makridakis2009], lawyers [@Koehler2002], and business managers [@Russo1992].
Thus, I would not put much confidence in the predictions by fantasy football pundits.
The most common pattern of experts' predictions is that they show overextremity, that is, their predictions have probability judgments that tend to be too extreme, as described in Section @sec-calibration.
Overextremity of experts' predictions reflects the [overprecision](#sec-cognitiveBiasesOverconfidence) type of [overconfidence bias](#sec-cognitiveBiasesOverconfidence).
The degree of confidence of a person's predictions is often not a good indicator of the accuracy of their predictions [and confidence and prediction accuracy are sometimes inversely associated; @Silver2012].
[Heuristics](#sec-heuristics) such as the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment), [cognitive biases](#sec-cognitiveBiases) such as [confirmation bias](#sec-cognitiveBiasesConfirmation) [@Hoch1985; @Koriat1980], [fallacies](#sec-fallacies) such as the [base rate fallacy](#sec-fallaciesBaseRate) [@Eddy1982; @Koehler2002] could contribute to overconfidence of predictions.
[Poorly calibrated](#sec-calibration) predictions are especially likely when the [base rate](#sec-baseRate) is very low (e.g., suicide) or when the [base rate](#sec-baseRate) is very high [@Koehler2002].

### 2025-07-20 machine-learning.qmd
20250717 - chapter overviews
20250720 - summary of prediction accuracy

@@

@@

# Prepare data for merging
#-todo: calculate years_of_experience
## Use common name for the same (gsis_id) ID variable

@@

@@

# Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
#-todo: remove redundant variables
playerListToMerge <- list(
  nfl_players %>% filter(!is.na(gsis_id)),
  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
)

>>>>

This chapter provides an overview of machine learning.

@@

#-todo:
#-calculate years_of_experience
#-use common name for the same (gsis_id) ID variable
```

@@

# Prepare data for merging

@@

#-todo:
#-remove redundant variables
```

@@

# Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
playerListToMerge <- list(
  nfl_players %>% filter(!is.na(gsis_id)),
  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
)
